---
title: "Working with XML and JSON in R"
author: "Coffy Andrews-Guo"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment

Three of my favorite books on 'Self Development' were selected this assignment. The assignment involves working with XML and JSON files with these criteria: _(1) At least one of the books should have more than one author. (2)For each book, include the title, authors, and two or three other attributes that you find interesting._

The three books and criteria was separately created in three files:_(1) HTML (using an html table), (2) XML, and (3) JSON formats (e.g. “books.html”, “books.xml”, and “books.json”). These files were created in a code editor (Notepad++).

All three file sources were loaded in R as separate data frames. In my observation all three sources when converted to a data frame have a similar structure and parsing (to character class) expression.


## Load Package
```{r}
library("textreadr")  # package to read HTML file
library("XML")        # package to read XML file
library("methods")    # also load the other required XML package
library("xml2")       # alternative package to read XML file
library("jsonlite")   # package to read JSON files
library("rvest")      # package to parse html table into a data frame

library("tidyverse")
library("dplyr")
```



## HTML File
```{r}
HTML_file <- read_html("https://raw.githubusercontent.com/candrewxs/D607_Wk7_XMLnJSON/main/D607%20Wk7_HTML.html")

print(HTML_file)
```

Convert data type to an R data frame
```{r}
HTML_df <- HTML_file %>%
  html_table() %>%
  as.data.frame()

HTML_df
```

Tidy data
```{r}
# transform the ISBN entries numerics to character to show entire length
HTML_df$ISBN <- as.character(HTML_df$ISBN)

HTML_df
```





## XML File 
```{r}
url_xml <- "https://raw.githubusercontent.com/candrewxs/D607_Wk7_XMLnJSON/main/D607%20Wk7_XML.xml"

print(url_xml)
```
Convert XML file and tidy data in R
```{r}
book_xml <- read_xml(url_xml)
book_xml <- xmlParse(book_xml) %>% xmlToDataFrame()

Books.Id <- c(1, 2, 3)      # define new column to add
book_xml2 <- cbind(Books.Id, book_xml) # add column called 'Books.Id'

book_xml2
```



## JSON File
```{r}
JSON_file <- fromJSON("https://raw.githubusercontent.com/candrewxs/D607_Wk7_XMLnJSON/main/D607%20W7_JSON.json")

print(JSON_file)
```

Convert JSON file and tidy _was unable to remove index from table_
```{r}
JSON_df <- as.data.frame(JSON_file)

colnames(JSON_df) <- c("Books.Id", "Title", "Authors", "Length", "Review", "ISBN")

JSON_df$Authors <- as.character(JSON_df$Authors)
JSON_df$Length <- as.character(JSON_df$Length)
JSON_df$Review <- as.character(JSON_df$Review)
JSON_df$ISBN <- as.character(JSON_df$ISBN)
JSON_df

```




